Testable normative theories

### 2020-05-08 simple.tex
(Probably) wrong path for clarifiying possible status of a prop

The set $\iphi \subseteq \Phi$ is the set of interest. 
When $(\prop, \text{possible}) \in \iphi$, $\prop$ is considered as a possibility in the deliberated judgment of $i$ (it is not rejected), and when $(\prop, \text{sure}) \in \iPhi$, $\prop$ is considered as a certainty in the deliberated judgment of $i$. 
Part of the content of $\Phi$ is known as a result of a correct normative and empirical analysis.

Given $\ar_0 \in \gargs$, let $c(\ar_0) = \bigcup \set{\args \subseteq \gargs \suchthat \ar_0 = \sum \args} \subseteq \allargs$ denote the \emph{components} of $\ar_0$.

Given $\phi \in {\gleadsto}(\allargs)$, an argument $\ar \in \allargs$ is \emph{in favor of $\phi$}, $\ar \in \gargs[\phi, 0]$, iff $\ar \in [\gbeatsinv]^{2k}({\gleadstoinv}(\phi))$ for some $k \in \N$, where $R^\alpha$ denotes $\alpha$ times the composition of the relation $R$ with itself. (Throughout the article, $\N$ includes zero.) 
Define $\gargs[\phi] = \set{\ar \suchthat c(\ar) \subseteq \gargs[\phi, 0]}$.
Define $\gargs = \cup_{\phi \in {\gleadstoinv}(\allargs)} \gargs[\phi]$.

@@

\subsection{Classify propositions}
The notations $\iphi$ and $\iphibar$ designate subsets of $\Phi$, representing the propositions that $i$ accepts and rejects, respectively.

\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$ 1]
	$\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \phi \in \iphi.
	\end{equation}
\end{axiom}
\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$ 2]
	$\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \lnot\phi \in \iphibar.
	\end{equation}
\end{axiom}
\begin{axiom}[Coherent complement]
	$\forall \phi \in \Phi: \phi \in \iphi ⇒ \phi \notin \iphibar$.
\end{axiom}

@@

\begin{theorem}
	Under axioms of Normative adequacy, Coherent complement ⇒ Coherent protocol.
\end{theorem}
\begin{proof}
	Coherent complement is equivalent to $\forall \phi \in \phi: ¬(\phi \in \iphi \land \phi \in \iphibar)$.
	Coherent protocol is equivalent to $\forall \phi \in \Phi: ¬(\ar \ileadstost \phi \land \ar \ileadstost ¬\phi)$, which implies (using Norm ad 1 and 2) $\forall \phi \in \Phi: ¬(\phi \in \iphi \land \phi \in \iphibar)$.
\end{proof}

\begin{theorem}
	Under axioms of Normative adequacy, Coherent protocol ⇒ Coherent complement.
\end{theorem}
\begin{proof}
	This is false under the current definitions: if $\ileadstost = \emptyset$, coherent protocol, but $\phi \in \iphi \land \phi \in \iphibar$ possible.
\end{proof}

\begin{axiom}[Coherent negation]
	$\forall \phi \in \Phi: \phi \in \iphi ⇔ ¬\phi \in \iphibar$.
\end{axiom}
This notation (writing $\phi \in \iphibar$ as equivalent to $\phi \notin \iphi$) is justified only if it is impossible that both $\phi \in \iphi$ and $\phi \notin \iphi$ be implied (thus, if Coherent complement). \commentOC{TODO clarify.} The following conditions permit to obtain this result. \commentOC{Cond. 1 is stronger than necessary: we could have ¬cond. 1 and still Coh compl: $(\ar_2, \ar_1) \ileadstost \phi$ and $(\ar_2, \ar_1) \ileadstost ¬\phi$ without problems if $\ar_2$ and $\ar_1$ are not decisive. In fact, $(\ar_2, \ar_1) \ileadstoe ¬\phi$ should be defined as $(\ar_2, \ar_1) \nileadstost \phi) \lor (\ar_2, \ar_1) \ileadstost ¬\phi$, just in case a strange protocol allows both leading always to $\phi$ and $¬\phi$, and I think the rest goes through.}

@@

\subsection{Validity and conditions}
\begin{definition}[Validity]
	A theory is \emph{valid} iff $\forall \phi \in {\gleadsto}(\allargs): [\phi \in \iphi \land ¬\phi \notin \iphi]$.
\end{definition}

@@

Define $\mbeats[g] \subseteq \gargs × \allargs$ as $\ar_2 \mbeats[g] \ar_1$ iff $\exists \ar \in c(\ar_2) \suchthat \ar \mbeats \ar_1$.
Define $\mbeats[a] \subseteq \allargs × \gargs$ as $\ar_1 \mbeats[a] \ar_0$ iff $\exists \ar \in c(\ar_0) \suchthat \ar_1 \mbeats \ar$ and $\nexists \ar \in c(\ar_0) \suchthat \ar \mbeats \ar_1$. Define ${\gbeatsstar} \subseteq \allargs × \allargs$ as ${\gbeatsstar} = {\mbeats[g]} \cup {\mbeats[a]}$.

@@

Given $\phi \in {\gleadsto}(\allargs)$, define ${\ibeatse[!\phi]} \subseteq \allargs × \gargs[\phi]$ such that, when $\phi = (\prop, \text{sure})$, $\ar' \ibeatse[!\phi] \ar$ iff $(\ar', \ar) \ileadstoe (¬\prop, \text{poss}) \lor (\ar', \ar) \ileadstoe (¬\prop, \text{sure})$, and when $\phi = (\prop, \text{poss})$, $\ar' \ibeatse[!\phi] \ar$ iff $(\ar', \ar) \ileadstoe ¬\phi$. Define ${\ibeatse[\phi]} \subseteq \gargs[\phi] × \allargs$ such that $\ar' \ibeatse[\phi] \ar$ iff $(\ar', \ar) \ileadstoe \phi$.

The following condition mandates that if the arguments used by the theory are not decisive, it knows a reason.
\begin{definition}[Answerability]
	$\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: (\ar_1, \ar_0) \ileadstoe ¬\phi ⇒ [\exists \ar_1' \in {\gbeatsstarinv}(\ar_0)]$.
\end{definition}

@@

\begin{definition}[Operational validity]
	$\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: [\ar_1 \gbeatsstar \ar_0 \land (\ar_1, \ar_0) \ileadstoe ¬\phi] ⇒ \exists \ar_2 \in {\gbeatsinv}(\ar_1)$.
\end{definition}

\begin{definition}[Claiming certainty requires claiming possibility]
	$\forall (\prop, \text{sure}) \in {\gleadsto}(\allargs): (\prop, \text{poss}) \in {\gleadsto}(\allargs)$.
\end{definition}

\begin{definition}[Challenging possibility implies challenging certainty]
	$\forall (\prop, \text{sure}) \in {\gleadsto}(\allargs), \ar_1 \in \allargs, \ar_0 \in \gargs[(\prop, \text{sure})]: (\ar_1, \ar_0) \ileadstoe (¬\prop, \text{sure}) ⇒ (\ar_1, \ar_0) \ileadstoe (¬\prop, \text{poss})$.
\end{definition}

>>>>

The set $\iPhi \subseteq \Phi$ is the set of interest. 
When $(\prop, \text{possible}) \in \iPhi$, $\prop$ is considered as a possibility in the deliberated judgment of $i$ (it is not rejected), and when $(\prop, \text{sure}) \in \iPhi$, $\prop$ is considered as a certainty in the deliberated judgment of $i$. 
Part of the content of $\Phi$ is known as a result of a correct normative and empirical analysis.

Given $\phi \in {\gleadsto}(\allargs)$, an argument $\ar \in \allargs$ is \emph{in favor of $\phi$}, $\ar \in \gargs[\phi, 1]$, iff $\ar \in [\gbeatsinv]^{2k}({\gleadstoinv}(\phi))$ for some $k \in \N$, where $R^\alpha$ denotes $\alpha$ times the composition of the relation $R$ with itself. (Throughout the article, $\N$ includes zero.)

Given $\phi \in {\gleadsto}(\allargs)$ and $1 ≤ k \in \N$, define $\gargs[\phi, k + 1] = \gargs[\phi, k] + \gargs[\phi, 1] = \set{\ar_k + \ar_1 \suchthat \ar_k \in \gargs[\phi, k], \ar_1  \in \gargs[\phi, 1]}$. 
Define $\gargs[\phi] = \bigcup_{1 ≤ k \in \N} \gargs[\phi, k]$, and $\gargs = \cup_{\phi \in {\gleadstoinv}(\allargs)} \gargs[\phi]$.

Given $\phi \in \Phi$ and $\ar_0 \in \gargs[\phi]$, let $c^\phi(\ar_0) = \bigcup \set{\args \subseteq \gargs[\phi, 0] \suchthat \ar_0 = \sum \args} \subseteq \allargs$ denote the \emph{components} of $\ar_0$ in favor of $\phi$.

@@

Given $\phi \in {\gleadsto}(\allargs)$, define ${\ibeatse[!\phi]} \subseteq \allargs × \gargs[\phi]$ such that $\ar' \ibeatse[!\phi] \ar$ iff $(\ar', \ar) \ileadstoe ¬\phi$. Define ${\ibeatse[\phi]} \subseteq \gargs[\phi] × \allargs$ such that $\ar' \ibeatse[\phi] \ar$ iff $(\ar', \ar) \ileadstoe \phi$.

\subsection{Classify propositions}
The notations $\iPhi$ and $\iPhirej$ designate the subsets of $\Phi$ whose propositions $i$ deliberately accepts and rejects, respectively.

\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$ 1]
	$\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \phi \in \iPhi.
	\end{equation}
\end{axiom}
\begin{axiom}[Normative adequacy of $\allargs$ and $\ileadstost$ 2]
	$\forall \phi \in \Phi$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \lnot\phi \in \iPhirej.
	\end{equation}
\end{axiom}
\begin{property}[Coherent complement]
	$\iPhi \cap \iPhirej = \emptyset$.
\end{property}

The following definition may be useful for analytic clarification, but this property is not required as an axiom for the results in this article.
\begin{property}[Exhaustivity]
	$\forall \phi \in \Phi: \phi \notin \iPhi ⇒ \phi \in \iPhirej$.
\end{property}
Exhaustivity can be considered a strong demand, as it requires every proposition to be deliberately accepted or rejected. \commentOC{I’m not so sure: could consider that $t$ is merely possible when not determined.}

Define the propositions whose status is determined as $\iPhidet = \iPhi \cup \iPhirej$.

\begin{property}[Suppl]
	$\forall \prop \in \allprops: \propsure \notin \iPhi ⇒ \propsure \in \iPhirej$.
\end{property}
\begin{property}[Negation coherence]
	\label{prop:cohb}
	$\forall \phi \in \Phi: \phi \in \iPhi ⇔ ¬\phi \in \iPhirej$, and $\forall \prop \in \allprops: \propsure \in \iPhi ⇒ (¬\prop, \text{sure}) \in \iPhirej$.
\end{property}
This property entails the following one.
\begin{property}[Per-proposition coherence]
	$\forall \prop \in \allprops$, exactly one of the following four propositions is true: 
	\begin{itemize}
		\item $\propsure \in \iPhi \land \propposs \in \iPhi \land (¬\prop, \text{sure}) \in \iPhirej \land (¬\prop, \text{poss}) \in \iPhirej$;
		\item $\propsure \in \iPhirej \land \propposs \in \iPhi \land (¬\prop, \text{sure}) \in \iPhirej \land (¬\prop, \text{poss}) \in \iPhi$;
		\item $\propsure \in \iPhirej \land \propposs \in \iPhirej \land (¬\prop, \text{sure}) \in \iPhi \land (¬\prop, \text{poss}) \in \iPhi$;
		\item $\set{\prop, ¬\prop} × \set{\text{sure}, \text{poss}} \subseteq \Phi \setminus \iPhidet$;
	\end{itemize}
\end{property}
\begin{proof}
	If $\propsure \in \iPhi$, then $\notpropsure \in \iPhirej$, thus $\propposs \in \iPhi$ as $\notpropsure = ¬\propposs$ and $\phi \in \iPhi ⇔ ¬\phi \in \iPhirej$; and finally, $\notpropposs \in \iPhirej$ as $\notpropposs = ¬\propsure$ and $\phi \in \iPhi ⇔ ¬\phi \in \iPhirej$.

	If $\propsure \notin \iPhi$ and $\propposs \in \iPhi$, then?
\end{proof}

This property in turn entails the following one.
\begin{property}[Per-proposition determination]
	$\forall \prop \in \allprops: [\set{\prop, ¬\prop} × \set{\text{sure}, \text{poss}} \subseteq \iPhidet] \lor [\set{\prop, ¬\prop} × \set{\text{sure}, \text{poss}} \subseteq \Phi \setminus \iPhidet]$.
\end{property}

Given a normative model, define $\mPhi \subseteq \Phi$ as the set of propositions whose status is determined by the Normative adequacy axioms: $\phi \in \mPhi ⇔ [\exists \ar \in \allargs \suchthat \ar \ileadstost \phi]$.

\Cref{prop:cohb} can’t be proven, but the following weaker versions can: $\forall \phi \in \Phi: \phi \in \mPhi ⇒ ¬\phi \in \iPhirej$ and $\forall \phi \in \Phi: \propsure \in \mPhi ⇒ ¬\propsure \in \iPhirej$.

\begin{property}[Per-proposition weaker coherence bis]
	$\forall \phi \in \Phi: \phi \in \mPhi ⇒ ¬\phi \notin \mPhi$.
\end{property}

A proposition $\phi$ has a determined status iff $\phi \in \mPhi \lor ¬\phi \in \mPhi$. (Not both?)

\begin{property}[Coherent determination]
	$\forall \prop \in \allprops: $ either all propositions among $\set{\propsure, \propposs, ¬\propsure, ¬\propposs}$ have a determined status, or none have.
\end{property}

… and it is coherent!

the subset of $\Phi$ about which $\allargs$ and $\ileadstoe$ give access:

@@

\begin{theorem}
	Under axioms of Normative adequacy, Coherent complement ⇒ Coherent protocol.
\end{theorem}
\begin{proof}
	Coherent complement is equivalent to $\forall \phi \in \phi: ¬(\phi \in \iPhi \land \phi \in \iPhirej)$.
	Coherent protocol is equivalent to $\forall \phi \in \Phi: ¬(\ar \ileadstost \phi \land \ar \ileadstost ¬\phi)$, which implies (using Norm ad 1 and 2) $\forall \phi \in \Phi: ¬(\phi \in \iPhi \land \phi \in \iPhirej)$.
\end{proof}

\begin{theorem}
	Under axioms of Normative adequacy, Coherent protocol ⇒ Coherent complement.
\end{theorem}
\begin{proof}
	This is false under the current definitions: if $\ileadstost = \emptyset$, coherent protocol, but $\phi \in \iPhi \land \phi \in \iPhirej$ possible.
\end{proof}

\begin{axiom}[Coherent negation]
	$\forall \phi \in \Phi: \phi \in \iPhi ⇔ ¬\phi \in \iPhirej$.
\end{axiom}

\begin{proposition}
	$\forall \ar \in \allargs: \ar \ileadstost \propsure ⇒ \ar \ileadstost \propposs$.
\end{proposition}
\begin{proof}
	By hypothesis, $\forall \ar_1 \in \allargs: (\ar_1, \ar) \ileadstost \propsure$, which implies, by definition of $\ileadstost$, that $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadstoe (¬\prop, \text{sure})$. Therefore, $\ar \ileadstost \propposs$.
\end{proof}

This notation (writing $\phi \in \iPhirej$ as equivalent to $\phi \notin \iPhi$) is justified only if it is impossible that both $\phi \in \iPhi$ and $\phi \notin \iPhi$ be implied (thus, if Coherent complement). \commentOC{TODO clarify.} The following conditions permit to obtain this result. \commentOC{Cond. 1 is stronger than necessary: we could have ¬cond. 1 and still Coh compl: $(\ar_2, \ar_1) \ileadstost \phi$ and $(\ar_2, \ar_1) \ileadstost ¬\phi$ without problems if $\ar_2$ and $\ar_1$ are not decisive. In fact, $(\ar_2, \ar_1) \ileadstoe ¬\phi$ should be defined as $(\ar_2, \ar_1) \nileadstost \phi) \lor (\ar_2, \ar_1) \ileadstost ¬\phi$, just in case a strange protocol allows both leading always to $\phi$ and $¬\phi$, and I think the rest goes through.}

@@

\subsection{Validity and conditions}
\begin{definition}[Validity]
	A theory is \emph{valid} iff $\forall \phi \in {\gleadsto}(\allargs): [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

@@

Define $\mbeats[g] \subseteq \gargs × \allargs$ as $\ar_2 \mbeats[g] \ar_1$ iff $\exists \ar \in c(\ar_2) \suchthat \ar \mbeats \ar_1$.
Define $\mbeats[a] \subseteq \allargs × \gargs$ as $\ar_1 \mbeats[a] \ar_0$ iff $\exists \ar \in c(\ar_0) \suchthat \ar_1 \mbeats \ar$ and $\nexists \ar \in c(\ar_0) \suchthat \ar \mbeats \ar_1$. Define ${\gbeatsstar} \subseteq \allargs × \allargs$ as ${\gbeatsstar} = {\mbeats[g]} \cup {\mbeats[a]}$. \commentOC{TODO $c^\phi$, not $c$. Use $c = \cup_{\phi} c^\phi$?}

@@

The following condition mandates that if the arguments used by the theory are not decisive, it knows a reason.
\begin{definition}[Answerability]
	$\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: (\ar_1, \ar_0) \nileadstost \phi ⇒ [\exists \ar_1' \in {\gbeatsstarinv}(\ar_0)]$.
\end{definition}

@@

\begin{definition}[Operational validity]
	$\forall \phi \in {\gleadsto}(\allargs), \ar_0 \in \gargs[\phi], \ar_1 \in \allargs: [\ar_1 \gbeatsstar \ar_0 \land (\ar_1, \ar_0) \nileadstost \phi] ⇒ \exists \ar_2 \in {\gbeatsinv}(\ar_1)$.
\end{definition}

\begin{definition}[Claiming certainty also claims possibility]
	$\forall \ar_0 \in \allargs: \ar_0 \gleadsto (\prop, \text{sure}) ⇒ \ar_0 \gleadsto (\prop, \text{poss})$.
\end{definition}

\begin{proposition}[Arguments for certainty also argue for possibility]
	From Claiming certainty also claims possibility follows that $\forall (\prop, \text{sure}) \in {\gleadsto}(\allargs): \ar_0 \in \gargs[(\prop, \text{sure})] ⇒ \ar_0 \in \gargs[(\prop, \text{poss})]$.
\end{proposition}
\begin{proof}
	TODO
\end{proof}

\begin{definition}
	$\forall \ar_0 \gleadsto (\prop, \text{sure}), \ar_1 \in \allargs: (\ar_1, \ar_0) \nileadstoe \phi ⇒ (\ar_1, \ar_0) \ileadstoe ¬\phi$.
\end{definition}
